---
title: "Personalized Marketing Campaign Analysis"
author: "Chunjian Zhang"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    theme: readable
  pdf:
    toc: true
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Load required libraries
library(tidyverse)
library(readr)
library(lubridate)
library(knitr)
library(kableExtra)
```

# Instruction

Personalized marketing is critical for driving customer engagement and sales in online retail. This report analyzes data from a recent quarter’s personalized marketing campaign. We examine how marketing spend per customer relates to purchase behavior. The goal is to provide data-driven insights for future marketing budget allocation.

# Research Question

Does higher marketing spend per customer lead to higher purchase amounts? We will analyse the relationship between marketing expenditure (`marketing_spend`) and the actual amount spent by customers (`purchase_amount`), whilst controlling for factors such as customer segment (`segment`) and number of visits (`visits`).

# **Dataset Introduction**

```{r data-cleaning, echo=FALSE, include=FALSE}
# Load raw data
df <- read.csv("data/35962712_Chunjian Zhang.csv")

# Clean date format
df <- df %>%
  mutate(
    # Convert mixed date formats to consistent Date type
    signup_date = case_when(
      str_detect(signup_date, "/") ~ dmy(signup_date),
      str_detect(signup_date, "-") ~ ymd(signup_date),
      str_detect(signup_date, ",") ~ mdy(signup_date),
      TRUE ~ NA_Date_
    ),
    # Create binary purchase indicator for conversion rate
    made_purchase = ifelse(!is.na(purchase_amount) & purchase_amount > 0, 1, 0)
  )

# Remove rows with missing purchase_amount for analysis
df_analysis <- df %>%
  filter(!is.na(purchase_amount))

# Create summary table for segment analysis
segment_summary <- df %>%
  group_by(segment) %>%
  summarise(
    n_customers = n(),
    avg_marketing_spend = mean(marketing_spend, na.rm = TRUE),
    avg_visits = mean(visits, na.rm = TRUE),
    avg_purchase = mean(purchase_amount, na.rm = TRUE),
    med_purchase = median(purchase_amount, na.rm = TRUE),
    conversion_rate = mean(made_purchase, na.rm = TRUE) * 100
  ) %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  bind_rows(
    df %>%
      summarise(
        segment = "Total",
        n_customers = n(),
        avg_marketing_spend = mean(marketing_spend, na.rm = TRUE),
        avg_visits = mean(visits, na.rm = TRUE),
        avg_purchase = mean(purchase_amount, na.rm = TRUE),
        med_purchase = median(purchase_amount, na.rm = TRUE),
        conversion_rate = mean(made_purchase, na.rm = TRUE) * 100
      ) %>%
      mutate(across(where(is.numeric), round, 2))
  )
```

This dataset comprises 150 customer observations and seven variables documenting customer identification, registration information, marketing expenditure, and transaction behavior. Core variables include `customer_id`, `marketing_spend`, `visits`, `purchase_amount`, and `segment`, which classifies customers into Consumer, SMB, and Enterprise groups. Data preprocessing standardized heterogeneous `signup_date` formats using the `lubridate` package to ensure temporal consistency. Observations with missing `purchase_amount` were excluded to support valid conversion rate computation, and duplicate `customer_id` entries were verified to be absent. Records with missing values in the non‑essential `notes` field were retained to preserve sample size and analytical power.
