---
title: "Personalized Marketing Campaign Analysis"
author: "Chunjian Zhang"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    theme: readable
  pdf:
    toc: true
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Load required libraries
library(tidyverse)
library(readr)
library(lubridate)
library(knitr)
library(kableExtra)
library(patchwork)
```

# Instruction

Personalized marketing is critical for driving customer engagement and sales in online retail. This report analyzes data from a recent quarter’s personalized marketing campaign. We examine how marketing spend per customer relates to purchase behavior. The goal is to provide data-driven insights for future marketing budget allocation.

# Research Question

Does higher marketing spend per customer lead to higher purchase amounts? We will analyse the relationship between marketing expenditure (`marketing_spend`) and the actual amount spent by customers (`purchase_amount`), whilst controlling for factors such as customer segment (`segment`) and number of visits (`visits`).

# **Dataset Introduction**

```{r data-cleaning, echo=FALSE, include=FALSE}
# Load raw data
df <- read.csv("data/35962712_Chunjian Zhang.csv")

# Clean date format
df <- df %>%
  mutate(
    # Convert mixed date formats to consistent Date type
    signup_date = case_when(
      str_detect(signup_date, "/") ~ dmy(signup_date),
      str_detect(signup_date, "-") ~ ymd(signup_date),
      str_detect(signup_date, ",") ~ mdy(signup_date),
      TRUE ~ NA_Date_
    ),
    # Create binary purchase indicator for conversion rate
    made_purchase = ifelse(!is.na(purchase_amount) & purchase_amount > 0, 1, 0)
  )

# Remove rows with missing purchase_amount for analysis
df_analysis <- df %>%
  filter(!is.na(purchase_amount))

# Create summary table for segment analysis
segment_summary <- df %>%
  group_by(segment) %>%
  summarise(
    n_customers = n(),
    avg_marketing_spend = mean(marketing_spend, na.rm = TRUE),
    avg_visits = mean(visits, na.rm = TRUE),
    avg_purchase = mean(purchase_amount, na.rm = TRUE),
    med_purchase = median(purchase_amount, na.rm = TRUE),
    conversion_rate = mean(made_purchase, na.rm = TRUE) * 100
  ) %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  bind_rows(
    df %>%
      summarise(
        segment = "Total",
        n_customers = n(),
        avg_marketing_spend = mean(marketing_spend, na.rm = TRUE),
        avg_visits = mean(visits, na.rm = TRUE),
        avg_purchase = mean(purchase_amount, na.rm = TRUE),
        med_purchase = median(purchase_amount, na.rm = TRUE),
        conversion_rate = mean(made_purchase, na.rm = TRUE) * 100
      ) %>%
      mutate(across(where(is.numeric), round, 2))
  )
```

This dataset comprises 150 customer observations and seven variables documenting customer identification, registration information, marketing expenditure, and transaction behavior. Core variables include `customer_id`, `marketing_spend`, `visits`, `purchase_amount`, and `segment`, which classifies customers into Consumer, SMB, and Enterprise groups. Data preprocessing standardized heterogeneous `signup_date` formats using the `lubridate` package to ensure temporal consistency. Observations with missing `purchase_amount` were excluded to support valid conversion rate computation, and duplicate `customer_id` entries were verified to be absent. Records with missing values in the non‑essential `notes` field were retained to preserve sample size and analytical power.

# **Dataset Description**

The cleaned dataset contains `r nrow(df)` observations and `r ncol(df)` variables. A brief summary of the data by customer segment is shown below.

```{r segment-table, echo=FALSE, warning=FALSE}
segment_summary <- df %>%
  group_by(segment) %>%
  summarise(
    n_customers = n(),
    avg_marketing_spend = mean(marketing_spend, na.rm = TRUE),
    avg_visits = mean(visits, na.rm = TRUE),
    avg_purchase = mean(purchase_amount, na.rm = TRUE),
    med_purchase = median(purchase_amount, na.rm = TRUE)
  ) %>%
  mutate(across(where(is.numeric), round, 2))

total <- df %>%
  summarise(
    segment = "Total",
    n_customers = n(),
    avg_marketing_spend = mean(marketing_spend, na.rm = TRUE),
    avg_visits = mean(visits, na.rm = TRUE),
    avg_purchase = mean(purchase_amount, na.rm = TRUE),
    med_purchase = median(purchase_amount, na.rm = TRUE)
  ) %>%
  mutate(across(where(is.numeric), round, 2))

final_table <- bind_rows(segment_summary, total)

final_table %>%
  select(
    Segment = segment,
    `Number of Customers` = n_customers,
    `Avg. Marketing Spend ($)` = avg_marketing_spend,
    `Avg. Visits` = avg_visits,
    `Avg. Purchase Amount ($)` = avg_purchase,
    `Med. Purchase Amount ($)` = med_purchase
  ) %>%
  kable(caption = "Key Metrics by Customer Segment") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
#| label: main-plot
#| echo: false
#| fig-cap: "Marketing Spend vs Purchase Amount and Average Visits by Customer Segment"
#| fig-width: 14
#| fig-height: 6
#| fig-dpi: 300

df_clean <- df %>% filter(!is.na(purchase_amount))

p1 <- ggplot(df_clean, aes(x = marketing_spend, y = purchase_amount, color = segment)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(x = "Marketing Spend ($)", y = "Purchase Amount ($)") +
  theme_minimal()

p2 <- df_clean %>%
  group_by(segment) %>%
  summarise(avg_visits = mean(visits)) %>%
  ggplot(aes(x = segment, y = avg_visits, fill = segment)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = round(avg_visits, 1)), vjust = -0.3) +
  labs(x = "Segment", y = "Avg Visits") +
  ylim(0, 7) +
  theme_minimal() +
  theme(legend.position = "none")

p1 + p2
ggsave("E:/University of Monash/ETC5513 Collaborative and Reproducible Practices/Assignment 1/assignment-1-creating-reproducible-reports-czha0262/images/customer_plot.png", plot = p1 + p2, width = 12, height = 5, dpi = 300)
```
